Supporting Information: Improving cropland maps through tight integration of human and machine intelligence

### 2020-08-10 inst/manuscript/figures.Rmd
Updated data analysis, figures, SI, manuscript through training data error impact section of results

### Figure 6: Categorical map accuracy

@@

### Figure S8: Segmentation statistics
```{r segstats, eval=FALSE}
data("top_label_data")
data("field_validation_stats")

>>>>

### Figure 6: The impact of training data error
```{r, eval=FALSE}
data("consensus_high_low")

V1 <- c("Accuracy", "AUC", "F1")

# calculate mean across AOIs and join. Convert to factor and reorder for
# plotting
consensus_high_low_mu <- consensus_high_low %>% 
  group_by(Strategy, Metric) %>% 
  summarize(Score = mean(Score)) %>% 
  ungroup() %>% 
  mutate(AOI = "Mean") %>% 
  bind_rows(consensus_high_low, .) %>% 
  filter(Metric %in% V1) %>% 
  mutate(AOI = factor(AOI)) %>% 
  mutate(AOI = forcats::fct_reorder(AOI, ord))

cols <- c("grey50", "grey60", "grey70", "grey80", "black")
sizes <- c(rep(1.5, 4), 2)
shapes <- c(3, 4, 8, 10, 16)
p <- ggplot(consensus_high_low_mu) + 
  geom_point(aes(x = Strategy, y = Score, color = AOI, size = AOI, 
                 shape = AOI)) + 
  ylim(0.4, 1) + xlab("Labelling strategy") + 
  scale_color_manual(values = cols) +
  scale_size_manual(values = sizes) + 
  scale_shape_manual(values = shapes) + 
  facet_wrap(~Metric) + 
  theme_linedraw() + 
  theme(legend.position = "bottom", 
        strip.background = element_rect(fill = "white"), 
        strip.text = element_text(colour = "black"))
ggsave(pth("figures/figure6.png"), height = 4, width = 6, dpi = 300)
```

### Figure 7: Categorical map accuracy

@@

### Figure S8: Training data error impact on probability images
```{r, eval = FALSE}
probpal <- viridis::viridis
cprob <- raster(here::here("external/prob/image_c294_r532.tif"))
hprob <- raster(here::here("external/prob/image_c294_r532_high.tif"))
lprob <- raster(here::here("external/prob/image_c294_r532_low.tif"))

png(pth("figures/si_label_strategy_probs.png"), height = 2.5,
    width = 8, res = 300, units = "in", bg = "transparent")
par(mfrow = c(1, 3), mar = c(0, 0.5, 0, 1.75), oma = c(0, 0, 0, 2))
plot(as(extent(cprob), "SpatialPolygons"), lty = 0)
plot(cprob, lty = 0, add = TRUE, legend = FALSE, zlim = c(0, 100),
     col = probpal(20))
mtext("A", side = 3, line = -2.5, adj = -0.02, cex = 1)
plot(as(extent(cprob), "SpatialPolygons"), lty = 0, cex.main = 2)
plot(hprob, lty = 0, add = TRUE, legend = FALSE, zlim = c(0, 100),
     col = probpal(20))
mtext("B", side = 3, line = -2.5, adj = -0.02, cex = 1)
plot(as(extent(cprob), "SpatialPolygons"), lty = 0)
plot(lprob, lty = 0, add = TRUE, zlim = c(0, 100),
     legend.width = 1, legend.shrink = 0.5, col = probpal(20))
mtext("C", side = 3, line = -2.5, adj = -0.02, cex = 1)
dev.off()

### Figure S9: Segmentation statistics
```{r segstats, eval=FALSE}
data("top_label_data")
data("field_validation_stats")

### 2020-08-10 inst/manuscript/manuscript.Rmd
Updated data analysis, figures, SI, manuscript through training data error impact section of results

#### Independent accuracy assessment
The accuracy assessment procedure described above was not strictly independent because it was re-used the same validation sites over multiple iterations [@ElmesAccountingtrainingdata2020]. To independently assess the accuracy of final map products, we followed guidelines provided by [@StehmanKeyissuesrigorous2019] to create a separate map reference sample. To develop this sample, we first used the segmented field boundaries to define a cropland and non-cropland stratum for Ghana, and then randomly selected a sample of the centroids of grid cells intersecting each stratum. Next, a primary supervisor examined the image composites in each randomly selected cell, and placed a square polygon of ~0.1 ha on the primary class intersecting the centerpoint of the cell. The square was shifted so that it was contained entirety with the intersected class: cropland or non-cropland. Where the supervisor was uncertain about the class label, one of two additional classes were selected: likely cropland or likely non-cropland. After placing these polygons, the supervisor removed their classes from the polygons and sent them to three separate experts, who labelled the polygons according to their interpretation of the underlying imagery. The four sets of labels were then combined and the overall agreement between them was calculated. This agreement provided the upper bound on knowable map accuracy. See the SI for further details on the creation and analysis of the map reference sample. These points were then intersected with both the RandomForest probability images and the segmented field boundaries, and confusion matrixes between the map reference sample labels and the extracted map classes was constructed to assess the categorical accuracy of each map product.

## The impact of training data error
To quantify the potential deleterious impact of training data error on our map results, we compared maps created by training the RandomForest algorithm with three sets of labels: 1) those from the lowest scoring worker to map each training site, 2) those from the most accurate worker, and 3) those trained as usual with the consensus labels. We then quantified the prediction performance against the validation sites, and compared the differences in performance metrics for each variant. This assessment follows recommended Tier 1 (i.e. best practice) standards to account for training data errors in machine learning-based Earth Observation applications [@ElmesAccountingtrainingdata2020].

@@

@@

## The characteristics of Ghanaian cropland
Using the final mapped results, we calculated the estimated area of arable cropland in Ghana, as well as the characteristics of field sizes through the country. We further used the uncertainty information encoded in the cropland probability images produced by the Random Forest classifier to provide further insight into the characteristics of cropland active-fallow cycles.

@@

@@

>>>>

#### Evaluating the impact of training data error
To quantify the potential impact of training data error on our map results, we compared maps created by training the RandomForest algorithm with three sets of labels: 1) those from the lowest scoring worker to map each training site, 2) those from the most accurate worker, and 3) those trained as usual with the consensus labels. We then quantified the prediction performance against the validation sites, and compared the differences in performance metrics for each variant. This assessment follows recommended Tier 1 (i.e. best practice) standards to account for training data errors in machine learning-based Earth Observation applications [@ElmesAccountingtrainingdata2020].

@@

#### Accuracy assessment
The accuracy assessment procedure described above was not strictly independent because it was re-used the same validation sites over multiple iterations [@ElmesAccountingtrainingdata2020]. To independently assess the accuracy of final map products, we followed guidelines provided by [@StehmanKeyissuesrigorous2019] to create a separate map reference sample. To develop this sample, we first used the segmented field boundaries to define a cropland and non-cropland stratum for Ghana, and then randomly selected a sample of the centroids of grid cells intersecting each stratum. Next, a primary supervisor examined the image composites in each randomly selected cell, and placed a square polygon of ~0.1 ha on the primary class intersecting the centerpoint of the cell. The square was shifted so that it was contained entirety with the intersected class: cropland or non-cropland. Where the supervisor was uncertain about the class label, one of two additional classes were selected: likely cropland or likely non-cropland. After placing these polygons, the supervisor removed their classes from the polygons and sent them to three separate experts, who labelled the polygons according to their interpretation of the underlying imagery. The four sets of labels were then combined and the overall agreement between them was calculated. This agreement provided the upper bound on knowable map accuracy. See the SI for further details on the creation and analysis of the map reference sample. These points were then intersected with both the RandomForest probability images and the segmented field boundaries, and confusion matrixes between the map reference sample labels and the extracted map classes was constructed to assess the categorical accuracy of each map product.

@@

## Assessing the characteristics of Ghanaian cropland
Using the final mapped results, we calculated the estimated area of arable cropland in Ghana, as well as the characteristics of field sizes through the country. We also used the uncertainty information encoded in the cropland probability images produced by the Random Forest classifier to provide further insight into the characteristics of cropland active-fallow cycles.

@@

### The impact of training data error
```{r, echo=FALSE, message=FALSE, warning=FALSE}
data("consensus_high_low")
V1 <- c("Accuracy", "AUC", "F1")
delta_func <- function(x, y, rnd = 1) round((x - y) / y * 100, rnd)

# calculate mean across AOIs and join. Convert to factor and reorder for
# plotting
consensus_high_low_mu <- consensus_high_low %>% 
  group_by(Strategy, Metric) %>% 
  summarize(Score = mean(Score)) %>% 
  ungroup()

metdiffs <- lapply(V1, function(x) {  # x <- "F1"
  cn <- consensus_high_low_mu %>% 
    filter(Strategy == "Consensus" & Metric == x) %>% pull(Score)
  h <- consensus_high_low_mu %>% filter(Strategy == "High" & Metric == x) %>% 
    pull(Score)
  l <- consensus_high_low_mu %>% filter(Strategy == "Low" & Metric == x) %>% 
    pull(Score)
  cnh <- delta_func(cn, h)
  cnl <- delta_func(cn, l)
  hl <- delta_func(h, l)
  list(cons = round(cn, 3), high = round(h, 3), low = round(l, 3), 
       chdelt = cnh, cldelt = cnl, hldelt = hl) 
})
names(metdiffs) <- V1
```

The potential impact of training data errors on map quality was assessed in four AOIs (1, 2, 8, and 15). The results of these tests showed that the average accuracy, AUC, and F1 scores for models trained with the consensus labels were respectively `r metdiffs$Accuracy$cons`, `r metdiffs$AUC$cons`, and `r metdiffs$F1$cons` (Figure \ref{fig:trainingimpact}). Models trained with the most accurate worker's labers had scores that were just `r metdiffs$AUC$chdelt` - `r metdiffs$Accuracy$chdelt` percent higher than models trained with the most accurate worker's lablers (accuracy = `r metdiffs$Accuracy$high`'; AUC = `r metdiffs$AUC$high`; F1 = `r metdiffs$F1$high`), but were `r metdiffs$AUC$cldelt` - `r metdiffs$Accuracy$cldelt` higher than models trained with the least accurate worker's labels (accuracy = `r metdiffs$Accuracy$low`; AUC = `r metdiffs$AUC$low`; F1 = `r metdiffs$F1$low`).

```{r trainingimpact, echo = FALSE, out.width="100%", fig.cap="Scores for overall Accuracy, area under the curve of the Receiver Operating Characteristic, and the F1 score resulting from Random Forest models trained with consensus labels, and labels made by the most and least accurate worker to label each site. Comparisons were made for AOIs 1, 2, 8, and 15, denoted by grey symbols, while the mean scores across these AOIs are shown for each metric.", fig.align='center', message=FALSE}
knitr::include_graphics('figures/figure6.png')
```

@@

- The impact of training data error on map quality:
  - Label quality makes a large difference. 
  - Although these differences were assessed against the validation sites, which were themselves consensus labels, there are real differences in quality that can be seen in the resulting probability images, where probabilities from the model trained with the least accurate worker's labels are clearly less certain (Figure S6).

### 2020-08-10 inst/manuscript/si.Rmd
Updated data analysis, figures, SI, manuscript through training data error impact section of results

@@

## Experiments
### Label quality

```{r, echo = FALSE, out.width="100%", fig.cap="Cropland probability images produced by RandomForest models trained with A) consensus labels, B) the most accurate worker's labels, and C) the least accurate worker's labels.", fig.align='center', message=FALSE}
knitr::include_graphics('figures/si_hml_probs.png')
```

>>>>

### The impact of training data error
Probability images resulting from Random Forest models trained with labels generated by 1) consensus and the 2) most and 3) least accurate worker to digitize each training site are illustrated in Figure \ref{fig:labelstrategy}. These images were created for a single tile in AOI 1.

```{r labelstrategy, echo = FALSE, out.width="100%", fig.cap="Cropland probability images produced by RandomForest models trained with A) consensus labels, B) the most accurate worker's labels, and C) the least accurate worker's labels.", fig.align='center', message=FALSE}
knitr::include_graphics('figures/si_label_strategy_probs.png')
```

@@
